---
title: "SSI_data_preprocessing"
output: html_document
date: "2022-11-08"
---

## loading necessary libraries
```{r echo=FALSE}
library(tidyverse)
library(data.table)
library(openxlsx)
```

## function to read data
```{r}
read_data<-function(URL)
{
 
  data <- fread(URL)%>%select(Year,County,Operative_Procedure,Facility_ID,Hospital_Category_RiskAdjustment,Procedure_Count,Infections_Reported)
  return(data)
  
}
```
## list of data URLs 
```{r}
URL_list <- c("https://data.chhs.ca.gov/dataset/f243090b-4c05-4c61-b614-7cb49b86b21d/resource/294eff96-096e-4b91-bff5-8de78fda318b/download/cdph_ssi_adult_odp_2021.csv",
              "https://data.chhs.ca.gov/dataset/f243090b-4c05-4c61-b614-7cb49b86b21d/resource/e6828e3c-bf54-4991-865b-b4cc11719b8d/download/cdph_ssi_adult_odp_2020h1.csv",
              "https://data.chhs.ca.gov/dataset/f243090b-4c05-4c61-b614-7cb49b86b21d/resource/eb324b4d-d2d2-41e8-aa53-f1b8380b4692/download/cdph_ssi_adult_odp_2020h2.csv",
              "https://data.chhs.ca.gov/dataset/f243090b-4c05-4c61-b614-7cb49b86b21d/resource/238ff746-71d9-4507-95d0-2a7b5146f4f5/download/cdph_ssi_adult_odp_2019.csv",
              "https://data.chhs.ca.gov/dataset/f243090b-4c05-4c61-b614-7cb49b86b21d/resource/bfbface5-a54a-43ae-a9a0-74d70dcc9d18/download/cdph_ssi_adult_odp_2018.csv"
              )
```
## applying the function to read all datasets
```{r}
SSI_adult_patients<- lapply(URL_list,read_data)
```
## changing list of lists to df format
```{r}
SSI_adult_patients_unzip<- do.call(rbind,SSI_adult_patients)
```
## check data types 
```{r}
str(SSI_adult_patients_unzip)
```
## change empty values with NAs
```{r}
SSI_adult_patients_unzip<-replace(SSI_adult_patients_unzip,SSI_adult_patients_unzip=="",NA)
```

## check number of missing values per attribute
```{r}
colSums(is.na(SSI_adult_patients_unzip))
```
## Removing missing values
```{r}
SSI_adult_patients_unzip_clean<-SSI_adult_patients_unzip[complete.cases(SSI_adult_patients_unzip)]
```
## Reading facility license dataset
```{r}
licensed_facilities = read.xlsx("https://data.chhs.ca.gov/dataset/e89100fd-1f1d-4a37-8205-d588aa42e5a1/resource/e083f267-43bf-427b-9ee3-f53faaff5d7e/download/odp_elms_aspen_oshpd_cw_mar2022.xlsx",sheet=1)
```

## change the column name to merge with the SSI dataset
```{r}
colnames(licensed_facilities)[colnames(licensed_facilities) == 'ELMS_FACID'] <- 'Facility_ID'
```
## change column type merge with the SSI dataset
```{r}
licensed_facilities$Facility_ID<-as.numeric(licensed_facilities$Facility_ID)
```
## join two datasets
```{r}
SSI_fac_ID_joined<-SSI_adult_patients_unzip_clean%>%inner_join(licensed_facilities,by="Facility_ID")
```
## Keep columns we need
```{r}
SSI_fac_ID_final<-SSI_fac_ID_joined%>% select(Year,County,Operative_Procedure,Hospital_Category_RiskAdjustment,Procedure_Count,Infections_Reported,LICENSED_CERTIFIED,FAC_TYPE_DESCRIPTION,FAC_FAC_RELATIONSHIP)
```
## check for missing values
```{r}
colSums(is.na(SSI_fac_ID_final))
```



